\documentclass{bschlangaul-aufgabe}
\bLadePakete{mathe,komplexitaetstheorie}
\begin{document}
\bAufgabenMetadaten{
  Titel = {Aufgabe 4},
  Thematik = {SAT DOPPELSAT},
  Referenz = 66115-2019-H.T1-A4,
  RelativerPfad = Staatsexamen/66115/2019/09/Thema-1/Aufgabe-4.tex,
  ZitatSchluessel = examen:66115:2019:09,
  BearbeitungsStand = mit Lösung,
  Korrektheit = unbekannt,
  Ueberprueft = {unbekannt},
  Stichwoerter = {Polynomialzeitreduktion},
  EinzelpruefungsNr = 66115,
  Jahr = 2019,
  Monat = 09,
  ThemaNr = 1,
  AufgabeNr = 4,
}

Betrachten Sie die folgenden Probleme:
\index{Polynomialzeitreduktion}
\footcite{examen:66115:2019:09}

\bProblemBeschreibung
{SAT}
{Aussagenlogische Formel F in KNF}
{Gibt es mindestens eine erfüllende Belegung für F?}

\bProblemBeschreibung
{DOPPELSAT}
{Aussagenlogische Formel F' in KNF}
{Gibt es mindestens eine erfüllende Belegung für F, in der
mindestens zwei Literale pro Klausel wahr sind?}

\begin{enumerate}

%%
% (a)
%%

\item Führen Sie eine polynomielle Reduktion von SAT auf DOPPELSAT durch.

\begin{bAntwort}
https://courses.cs.washington.edu/courses/csep531/09wi/handouts/sol4.pdf

DOUBLE-SAT is in NP. The polynomial size certificate consists of two
assignments f1 and f2. First, the verifier verifies if f1 6= f2. Then,
it verifies if both assignments satisfy φ by subtituting the values for
the variables and evaluate the clauses of φ. Both checks can be done in
linear time. DOUBLE-SAT is NP-hard. We give a reduction from SAT. Given
an instance φ of SAT which is a CNF formula of n variables x1,x2,...xn,
we construct a new variable xn+1 and let ψ = φ ∧(xn+1 ∨¬xn+1) be the
corresponding instance of DOUBLE-SAT. We claim that φ has a satisfying
assignment iff ψ has at least two satisfying assignments. On one hand,
if φ has a satisfying assignment f, we can obtain two distinct
satisfying assignments of ψ by extending f with xn+1 = T and xn+1 = F
respectively. On the other hand, if ψ has at least two sastisyfing
assignments then the restriction of any of them to the set {x1,x2,...xn}
is a satisfying assignment for φ. Thus, DOUBLE-SAT is NP-complete.

https://cs.stackexchange.com/questions/6371/proving-double-sat-is-np-complete

Here is one solution:

Clearly Double-SAT belongs to ${\sf NP}$, since a NTM can decide
Double-SAT as follows: On a Boolean input formula
$\phi(x_1,\ldots,x_n)$, nondeterministically guess 2 assignments and
verify whether both satisfy $\phi$.

To show that Double-SAT is ${\sf NP}$-Complete, we give a reduction from
SAT to Double-SAT, as follows:

On input $\phi(x_1,\ldots,x_n)$:

1. Introduce a new variable $y$.
2. Output formula $\phi'(x_1,\ldots,x_n, y) = \phi(x_1,\ldots,x_n) \wedge (y \vee \bar y)$.

If $\phi (x_1,\ldots,x_n)$ belongs to SAT, then $\phi$ has at least 1
satisfying assignment, and therefore $\phi'(x_1,\ldots,x_n, y)$ has at
least 2 satisfying assignments as we can satisfy the new clause ($y \vee
\bar y$) by assigning either $y = 1$ or $y = 0$ to the new variable $y$,
so $\phi'$($x_1$, ... ,$x_n$, $y$) $\in$ Double-SAT.

 On the other hand, if $\phi(x_1,\ldots,x_n)\notin \text{SAT}$, then
 clearly $\phi' (x_1,\ldots,x_n, y) = \phi (x_1,\ldots,x_n) \wedge (y
 \vee \bar y)$ has no satisfying assignment either, so
 $\phi'(x_1,\ldots,x_n,y) \notin  \text{Double-SAT}$.

Therefore, $\text{SAT} \leq_p \text{Double-SAT}$, and hence Double-SAT
is ${\sf NP}$-Complete.

\end{bAntwort}

%%
% (b)
%%

\item Zeigen Sie, dass DOPPELSAT NP-vollständig ist.

\end{enumerate}
\end{document}
